---
title: "Data Exercise - Bank Loan Defaults"
output:
  html_document:
    code_folding: show
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
subtitle: ExamPA.net
---

The first 10-20 minutes of the exam will be spent just on data manipulation.  This will count for between 15-20% of the final grade.  This notebook provides an example of how to do this.

## Data 

https://archive.ics.uci.edu/ml/datasets/Bank+Marketing

|Variable                                                 |Description                                                                                                                                         |Values                                                                                                                                                                                                                                                                                                                                                |
|---------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| age (numeric)                                           |                                                                                                                                                    |                                                                                                                                                                                                                                                                                                                                                      |
| job                                                     | type of job (categorical                                                                                                                           | 'admin.','blue                                                                                                                                                                                                                                                                                                                                       |
| marital                                                 | marital status (categorical                                                                                                                        | 'divorced','married','single','unknown'; note -  'divorced' means divorced or widowed)                                                                                                                                                                                                                                                               |
| education (categorical                                  | 'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')                                   |                                                                                                                                                                                                                                                                                                                                                      |
| default                                                 | has credit in default? (categorical                                                                                                                | 'no','yes','unknown')                                                                                                                                                                                                                                                                                                                                |
| housing                                                 | has housing loan? (categorical                                                                                                                     | 'no','yes','unknown')                                                                                                                                                                                                                                                                                                                                |
| loan                                                    | has personal loan? (categorical                                                                                                                    | 'no','yes','unknown')                                                                                                                                                                                                                                                                                                                                |
| contact                                                 | contact communication type (categorical                                                                                                            | 'cellular','telephone')                                                                                                                                                                                                                                                                                                                              |
| month                                                   | last contact month of year (categorical                                                                                                            | 'jan', 'feb', 'mar', ..., 'nov', 'dec')                                                                                                                                                                                                                                                                                                              |
| day_of_week                                             | last contact day of the week (categorical                                                                                                          | 'mon','tue','wed','thu','fri')                                                                                                                                                                                                                                                                                                                       |
| duration                                                | last contact duration, in seconds (numeric). Important note -  this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.|                                                                                                                                                                                                                                                                                                                                                      |
| campaign                                                | number of contacts performed during this campaign and for this client (numeric, includes last contact)                                             |                                                                                                                                                                                                                                                                                                                                                      |
| pdays                                                   | number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)|                                                                                                                                                                                                                                                                                                                                                      |
| previous                                                | number of contacts performed before this campaign and for this client (numeric)                                                                    |                                                                                                                                                                                                                                                                                                                                                      |
| poutcome                                                | outcome of the previous marketing campaign (categorical                                                                                            | 'failure','nonexistent','success')                                                                                                                                                                                                                                                                                                                   |
| emp.var.rate                                            | employment variation rate                                                                                                                          |                                                                                                                                                                                                                                                                                                                                                      |
| cons.price.idx                                          | consumer price index                                                                                                                               |                                                                                                                                                                                                                                                                                                                                                      |
| cons.conf.idx                                           | consumer confidence index                                                                                                                          |                                                                                                                                                                                                                                                                                                                                                      |
| euribor3m                                               | euribor 3 month rate                                                                                                                               |                                                                                                                                                                                                                                                                                                                                                      |
| nr.employed                                             | number of employees                                                                                                                                |                                                                                                                                                                                                                                                                                                                                                      |
| y                                                       |has the client subscribed a term deposit?                                                                                                           |yes', 'no'                                                                                                                                                                                                                                                                                                                                            |



```{r message = F}
library(tidyverse)
theme_set(theme_bw())
bank <- read_delim("bank-additional-full.csv", delim = ";")

```

## Task 1 - Examine the Target Varible

Examine the target variable y, which is whether or not the client has subscribed to a term deposite.

21 - y - has the client subscribed a term deposit? (binary: 'yes','no')

```{r}
bank %>% count(y)
bank %>% group_by(day_of_week) %>% summarise(percent_subscribed = mean(y=="yes"))
bank %>% group_by(education) %>% summarise(percent_subscribed = mean(y=="yes")) %>% arrange(desc(percent_subscribed))
```

## Task 2 - Decide on which variables to discard

Read the data dictionary and if appropriate, discard any variables that would not be useful for building a model.

```{r}
#the data dictionary says to not use duration in predictive modeling
bank <- bank %>% select(-duration)
```

We exclude the duration variable because it will not be known in advance of calling the individual and so should not be used in the predictive model.

"Yet, the duration is not known before a call is performed"

## Task 3 - Examine the numeric variables

- age 
- campaign 
- pdays 
- previous
- emp.var.rate, cons.price.idx, cons.conf.idx, euribor3m, nr.employed 

```{r}
bank %>% select_if(is.numeric) %>% summary()
bank %>% count(cons.price.idx)

#example creating a histogram
bank %>% ggplot(aes(cons.price.idx)) + geom_histogram()
bank %>% count(emp.var.rate)

bank %>% count(pdays)

#example of how to create bins
bank <- bank %>% 
  mutate(pdays_flag = ifelse(pdays == 999, yes = "Flag", "NoFlag"),
         pdays_bin = case_when(pdays <= 5 ~ "0-5",
                               pdays <= 10 ~ "0-10",
                               pdays <= 15 ~ "10-15",
                               pdays <= 20 ~ "15-20",
                               pdays <= 30 ~ "20-30", 
                               pdays == 999 ~ "None"))

#set factor levels to those with the most observations
bank <- bank %>% 
  mutate(pdays_flag = fct_infreq(pdays_flag),
         pdays_bin = fct_infreq(pdays_bin))

bank %>% count(pdays_flag)
bank %>% count(pdays_bin)


```

## Task 4 - Examine the factor variables

```{r}
bank %>% 
  mutate_if(is.character, as.factor) %>% 
  select_if(is.factor) %>% 
  summary()

fct_order <- bank %>% 
  count(job) %>% 
  arrange(n) %>% 
  select(job) %>% 
  unlist() %>% 
  as.character()

bank %>% 
  mutate(job = fct_relevel(job, fct_order)) %>% 
  ggplot(aes(job, fill = y)) + 
  geom_bar(stat = "count") + 
  coord_flip()

bank %>% 
  mutate(job = fct_relevel(job, fct_order)) %>% 
  ggplot(aes(job, fill = y)) + 
  geom_bar(stat = "count", position = "fill") + 
  coord_flip()

bank %>% 
  group_by(marital) %>% 
  summarise(percent_subscribed = round(mean(y == "yes"),2)) %>% 
  arrange(desc(percent_subscribed))
```

```{r}
fct_order <- bank %>% 
  count(marital) %>% 
  arrange(n) %>% 
  select(marital) %>% 
  unlist() %>% 
  as.character()

bank %>% 
  group_by(marital) %>% 
  summarise(percent_subscribed = mean(y == "yes"))

bank %>% 
  group_by(day_of_week, education) %>% 
  summarise(percent_subscribed = round(mean(y == "yes"),1))

bank %>% 
  count(job) %>% 
  arrange(desc(n)) %>% 
  mutate(pct = round(n/sum(n),1))

bank %>% 
  group_by(day_of_week, job, education, contact) %>% 
  summarise(percent_subscribed = round(mean(y == "yes"),2)) %>% 
  arrange(desc(percent_subscribed))
```

